---
title: "rapm_validator.Rmd"
output: html_document
date: "2023-12-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries.
```{r, echo=FALSE}
library(tidyverse)
library(glmnet)
library(data.table)
library(ggplot2)
```

Load the data, .csv files, from data folder. Set data to backup. 
```{r}
file_path <- "../data/input/pbp_events_L1_S2022_STRegular_fixed_coords.csv"

# Extract "L1_S20XX" using regular expression
season_info <- sub(".*/pbp_events_(L\\d+_S\\d+)_STRegular_fixed_coords.csv", "\\1", file_path)
```

assign shifts_grouped_ev table from calculator script
```{r}
shifts_grouped_ev = paste0("shifts_grouped_ev_", season_info)
assign(shifts_grouped_ev, readRDS(paste0("../data/temp/shifts_grouped_ev_", season_info, ".RData")))
shifts_grouped_ev = get(shifts_grouped_ev)
```

Create final table of the 21/22 season
```{r}
team_names_L1_21 <- c("EVZ", "HCL", "ZSC", "FRI", "LHC", "GSHC", "EHCB", "HCD", "SCB", "SCRJ", "HCAP", "SCL") # 2021
team_names_L1_22 <- c("EVZ", "FRI", "ZSC", "SCRJ", "HCD", "EHCB", "LHC", "GSHC", "HCL", "HCAP", "SCB", "SCL", "AJO") # 2022
team_names_L2_21 <- c("EHCK", "AJO", "LAN", "HCS", "OLT", "THU", "VIS", "HCC", "GCK", "rockets", "EVZA", "WIN")
team_names_L2_22 <- c("EHCK", "OLT", "HCC", "THU", "LAN", "VIS", "HCS", "GCK", "EVZA", "WIN", "rockets")

final_positions <- seq(along = team_names_L1_22)

final_table <- data.frame(Team = team_names_L1_22, Endplatzierung = final_positions)
```

Compute correlation between the average xGMP_60 value per team and the final table of the season
```{r}
rapm_with_team <- read_csv(paste0("../data/output/RapmEvenStrengthJoined_", season_info, ".csv"))

ranked_table <- rapm_with_team %>%
  group_by(Team) %>%
  mutate(rank = dense_rank(desc(xGPM_60)))

team_sum <- ranked_table %>%
  group_by(Team) %>%
  summarise(Avg_xGPM_60_Player = mean(xGPM_60, na.rm = TRUE))

team_sum_final_table <- arrange(team_sum, desc(Avg_xGPM_60_Player))
team_sum_final_table$Endplatzierung <- rank(desc(team_sum_final_table$Avg_xGPM_60_Player))

merged_table <- merge(team_sum_final_table, final_table, by = "Team")

correlation <- cor(merged_table$Endplatzierung.x, merged_table$Endplatzierung.y, method = "spearman")

print(correlation)
```

Korrelation zwischen den beiden Saisons nach player rank
```{r}
players_22 <- read_csv("../data/output/RapmEvenStrengthJoined_L1_S2022.csv")
players_21 <- read_csv("../data/output/RapmEvenStrengthJoined_L1_S2021.csv")

players_21$players_rank <- rank(desc(players_21$xGPM_60), na.last = "keep")
players_22$players_rank <- rank(desc(players_22$xGPM_60), na.last = "keep")

merged_table <- merge(players_21, players_22, by = c("Player", "Team"))
merged_table <- merged_table[merged_table$TimeOnIce.x >= 100 & merged_table$TimeOnIce.y >= 100, ]

correlation <- cor(merged_table$players_rank.x, merged_table$players_rank.y, method = "spearman")
print(correlation)
```

Korrelation zwischen den beiden Saisons nach xGPM_60 Wert der Spieler mit plot
```{r}
players_22 <- read_csv("../data/output/RapmEvenStrengthJoined_L2_S2022.csv")
players_21 <- read_csv("../data/output/RapmEvenStrengthJoined_L2_S2021.csv")

merged_table <- merge(players_21, players_22, by = c("Player", "Team")) # Without team also possible
merged_table <- merged_table[merged_table$TimeOnIce.x >= 100 & merged_table$TimeOnIce.y >= 100, ]

correlation <- cor(merged_table$xGPM_60.x, merged_table$xGPM_60.y, method = "pearson")
print(correlation)

p <- ggplot(merged_table, aes(x = xGPM_60.x, y = xGPM_60.y)) +
  geom_point(size = 3, color = "black") +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "blue", linewidth = 1) +
  geom_smooth(method = "lm", color = "gray", se = FALSE, linewidth = 0.5) +
  labs(
    x = "xGPM60 2021",
    y = "xGPM60 2022",
    title = "Korrelation der xGPM60-Werte über zwei Saisons"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  ) +
  guides(color = FALSE) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

print(p)
```

compute average shift lengths
```{r}
max_values <- tail(sort(shifts_grouped_ev$shift_length), 10)

hist(shifts_grouped_ev$shift_length, breaks = 50, main = "Die Schichtlängen in der National League", xlab = "Schichtlänge [s]", col = "powderblue")
mean_shift <- mean(na.omit(shifts_grouped_ev$shift_length))
median_shift <- median(na.omit(shifts_grouped_ev$shift_length))

abline(v = mean_shift, col = "red", lwd = 2)
abline(v = median_shift, col = "pink", lwd = 2)
 
text(x = 90, y = 20000, labels = paste("Median:", round(median_shift, 2)), col = "pink")
text(x = 90, y = 25000, labels = paste("Mittelwert:", round(mean_shift, 2)), col = "red")

```

Calculate the summed up xG from all players while they where on the field
```{r}
shifts_long <- shifts_grouped_ev %>%
  gather(key = "Player_position", value = "Player", h1.num:h6.num, a1.num:a6.num) %>%
  mutate(
    xG = ifelse(grepl("^h", Player_position), as.numeric(homexGF), as.numeric(awayxGF)),
    IsGoalie = (Player == Home.Goalie) | (Player == Away.Goalie)
  ) %>%
  filter(!IsGoalie, Player != "MISSING_PLAYER") # Herausfiltern der Torhüter und "MISSING_PLAYER"

xG_per_player <- shifts_long %>%
  group_by(Player) %>%
  summarise(TotalxG = sum(xG, na.rm = TRUE)) %>%
  arrange(desc(TotalxG))
```

Caculate the time on ice from all players
```{r}
# Reshape the data to have one row per player per shift
toi_data <- shifts_grouped_ev %>%
  select(h1.num:h6.num,a1.num:a6.num,shift_length)

toi_data <- toi_data %>%
  gather(key = "Player_position", value = "Player", h1.num:h6.num, a1.num:a6.num) %>%
  select(Player, shift_length)

# Calculate TOI for each player
toi <- toi_data %>%
  group_by(Player) %>%
  summarise(TimeOnIce = sum(shift_length, na.rm = TRUE) / 60) %>% # from seconds -> minutes
  arrange(desc(TimeOnIce))
```

calculate xG values for all players per 60
```{r}
actual_xG60 <- left_join(xG_per_player, toi, by = "Player")
actual_xG60 <- actual_xG60 %>%
  mutate(actual_xG_60 = (TotalxG / TimeOnIce) * 60)

write_excel_csv(actual_xG60, paste0("../data/temp/actual_xG60_per_Player_", season_info, ".csv"))
```

calculate correlation between xGF_60 and actual_xG60_per_player
```{r}
final_player_stats <- read_csv(paste0("../data/output/RapmEvenStrengthJoined_", season_info, ".csv"))
actual_xG60_per_player <- read_csv(paste0("../data/temp/actual_xG60_per_Player_", season_info, ".csv"))

final_player_stats_filtered <- final_player_stats %>%
  filter(TimeOnIce >= 100) %>%
  arrange(desc(xGF_60))

actual_xG60_per_player_filtered <- actual_xG60_per_player %>%
  filter(TimeOnIce >= 100) %>%
  arrange(desc(actual_xG_60))

combined_data <- merge(final_player_stats_filtered, actual_xG60_per_player_filtered, by = "Player")

pearson_corr <- cor(combined_data$xGF_60, combined_data$actual_xG_60, method = "pearson")

spearman_corr <- cor(combined_data$xGF_60, combined_data$actual_xG_60, method = "spearman")

p <- ggplot(combined_data, aes(x = xGF_60, y = actual_xG_60)) +
  geom_point(size = 3, color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(
    x = "xGF pro 60 Minuten",
    y = "Tatsächliche xG pro 60 Minuten",
    title = "Korrelation zwischen xGF und tatsächlichen xG"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, hjust = 0.5),
    legend.position = "none",  # Legende entfernen
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )

print(p)

print(spearman_corr)
```

```{r}
# Lets Run all the datasets to make some awesome plots:
# Define the folder path
folder_path <- "../data/input/"

# Define the regular expression pattern to match file names
pattern <- "pbp_events_(L\\d+_S\\d+)_STRegular_fixed_coords.csv"

# Get a list of file names that match the pattern in the folder
file_list <- list.files(path = folder_path, pattern = pattern, full.names = TRUE)
print(file_list)

# Take a bath or a nice lunchbreak!
for (file_path in file_list) {

# Set the file_path variable in the environment of the rapm_calculator.R script
  assign("file_path", file_path, envir = environment())
  # Run the rapm_calculator.R script
  source("rapm_calculator.R")
}
```

```{r}
library(ggplot2)
# Lets check for multiple seasons, leagues if our ridge regression is doing what she intends to!
# Filter defenders with less than 500 minutes of ToI
filtered_defenders <- rbind(filtered_defenders_L1_S2021,filtered_defenders_L1_S2022,filtered_defenders_L2_S2021,filtered_defenders_L2_S2022)

# Filter strikers with less than 500 minutes of ToI
filtered_striker <- rbind(filtered_strikers_L1_S2021,filtered_strikers_L1_S2022,filtered_strikers_L2_S2021,filtered_strikers_L2_S2022)

# Create scatter plot for defenders
plot_defenders_xGA <- ggplot(data = filtered_defenders, aes(x = TimeOnIce, y = xGA_60)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "steelblue") +
  ylim(-1, 1) +
  labs(x = "Eiszeit [min]", y = "xGA_60") +
  theme_minimal()
# Create scatter plot for strikers

plot_strikers_xGA <- ggplot(data = filtered_strikers, aes(x = TimeOnIce, y = xGA_60)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "hotpink") +
  ylim(-1, 1) +
  labs(x = "Eiszeit [min]", y = "xGA_60") +
  theme_minimal()

plot_defenders_xGF <- ggplot(data = filtered_defenders, aes(x = TimeOnIce, y = xGF_60)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "steelblue") +
  ylim(-1, 1) +
  labs(x = "Eiszeit [min]", y = "xGF_60") +
  theme_minimal()

# Create scatter plot for strikers
plot_strikers_xGF <- ggplot(data = filtered_strikers, aes(x = TimeOnIce, y = xGF_60)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "hotpink") +
  ylim(-1, 1) +
  labs(x = "Eiszeit [min]", y = "xGF_60") +
  theme_minimal()

plot_defenders_xGPM_60 <- ggplot(data = filtered_defenders, aes(x = TimeOnIce, y = xGPM_60)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "steelblue") +
  ylim(-1, 1) +
  labs(x = "Eiszeit [min]", y = "xGPM_60") +
  theme_minimal()

# Create scatter plot for strikers
plot_strikers_xGPM_60 <- ggplot(data = filtered_strikers, aes(x = TimeOnIce, y = xGPM_60)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "hotpink") +
  ylim(-1, 1) +
  labs(x = "Eiszeit [min]", y = "xGPM_60") +
  theme_minimal() 

library(gridExtra)
grid_plot <- grid.arrange(plot_defenders_xGF, plot_strikers_xGF, plot_defenders_xGA, plot_strikers_xGA,plot_defenders_xGPM_60,plot_strikers_xGPM_60, ncol = 2, nrow = 3)

ggsave("../data/output/xGF_60_ridge.png", grid_plot)
```

```{r}
# Create scatter plot for defenders
ggplot(data = rapm_defender, aes(x = TimeOnIce, y = xGA_60)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "steelblue") +
  ylim(-1, 1) +
  labs(x = "Time On Ice", y = "xGA_60") +
  theme_minimal()
which.max(rapm_defender$xGA_60)
rapm_defender[146,] # Helfer Even, spielte bei AJO 23 Partien und erhielt eine unmenge an TOREN!
```

lets do the same plot for the linear regression
```{r}
filtered_unique_500 <- filtered_unique[filtered_unique$TimeOnIce < 500, ]
filtered_unique_500 <- na.omit(filtered_unique_500)
filtered_unique_500

ggplot(data = filtered_unique_500, aes(x = TimeOnIce, y = values.Estimate)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "tomato") +
  ylim(-2.5, 2.5) +
  labs(x = "Eiszeit [min]", y = "xGF_60") +
  theme_minimal()
```

# Alle Datensätze für lineare Regression vorbereiten:
```{r}
# Define the directory where RData files are located
data_dir <- "../data/temp"  # Update with your directory path

# List of RData files to process
file_list <- list.files(path = data_dir, pattern = "^RAPM_xGF_L\\d{1}\\_S\\d{4}\\.RData$", full.names = TRUE)
print(file_list)

# Iterate over the list of file paths and process each file
for (temp_file_path in file_list) {
  print(temp_file_path)
  assign("temp_file_path", temp_file_path, envir = environment())
  # Run the rapm_calculator.R script
  source("linear_regression_validation.R")
}
```


```{r}
filtered_linear <- rbind(filtered_linear_L1_S2021,filtered_linear_L1_S2022,filtered_linear_L2_S2021,filtered_linear_L2_S2022)
filtered_linear <- filtered_linear[filtered_linear$TimeOnIce < 500, ]
filtered_linear <- na.omit(filtered_linear)

# Create scatter plot with legend
ggplot(data = filtered_linear, aes(x = TimeOnIce, y = values.Estimate)) +
  geom_point(shape = 21, size = 3, alpha = 0.3, fill = "tomato") +
  ylim(-2.5, 2.5) +
  labs(x = "Eiszeit [min]", y = "xGF_60 Lineare Regression") +
  theme_minimal()
ggsave("../data/output/xGF_60_lin.png")
```